Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-18 paper.tex
fix depth

With all of this in mind, in Fig.~\ref{fig:erase}, the test is set up as follows: given some rule $r$ with $n$ premises $P_i(x)$, $1 \leq i \leq n$, we run the procedure in a context where $n-1$ premises are instanciated with $a$ and one of them with $a + 1$.
The numbers on the $x$ axis represent which hypothesis has the incompatible instantiation in the order of consideration by the algorithm.
In a sense, this is how \textit{deep} the naive algorithm investigate before it stops.
The reason we see similar times for the naive algorithm at $1$ and $2$ is because for $1$, it tries to build the theorem using $P_n(a + 1)$ and immediately fails because $P_{n - 1}(a)$ is incompatible.
For $2$, the same thing happens but the instantiations are swapped.
For the next depths, the stateless algorithm gets progressively slower the further the incompatible hypothesis is located.
The position of the incompatible hypothesis has little effect on the stateful algorithm as the index selects the rule in all case.
The work the stateful algorithm has to do is thus close to constant in this scenario.
The Figure \ref{fig:erase-noprecomp} showcases that the naive algorithm has a sizable advantage in this test, especially when unification is difficult.

@@

\begin{figure}
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.75]
      \begin{axis}[
        title={Depth Test},
        xlabel={Depth},
        ylabel={Time in ms},
        xmin=1, xmax=6,
        ymin=0, ymax=2^10,
        xtick={1,2,3,4,5,6},
        ytick={4,16,64,256,1024},
        ymode=log,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 32.580083) (2, 33.212805) (3, 40.239833) (4, 47.670277) (5, 54.647458) (6, 61.044958) 
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 127.925541) (2, 140.497597) (3, 123.719264) (4, 130.819361) (5, 121.047667) (6, 123.789500)
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 108.120444) (2, 107.352694) (3, 116.182833) (4, 125.995639) (5, 137.815527) (6, 142.310597) 
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 604.595791) (2, 624.541194) (3, 608.671111) (4, 614.285014) (5, 588.636819) (6, 595.276806) 
          };
      \end{axis}
    \end{tikzpicture}
    \caption{Depth benchmark no-precompilation}%
    \label{fig:erase-noprecomp}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.75]
      \begin{axis}[
        title={Depth Test},
        xlabel={Depth},
        %ylabel={Time in ms},
        xmin=1, xmax=6,
        ymin=0, ymax=170,
        xtick={1,2,3,4,5,6},
        ytick={2^1,2^3,2^5,2^7},
        ymode=log,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 12.820514) (2, 12.645666) (3, 17.582138) (4, 22.730528) (5, 27.280097) (6, 32.578722) 
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 13.406416) (2, 13.747139) (3, 12.503764) (4, 12.950486) (5, 12.282083) (6, 12.151930) 
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 83.032819) (2, 79.396889) (3, 85.939180) (4, 92.495542) (5, 96.461291) (6, 105.517458)
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 25.681583) (2, 25.322444) (3, 27.672833) (4, 26.999055) (5, 25.690319) (6, 25.257291)
          };
      \end{axis}
    \end{tikzpicture}
    % graph \includegraphics[width=.4\linewidth]{image1}
    \caption{Depth benchmark with precompilation}%
    \label{fig:erase-precomp}
  \end{subfigure}
  \caption{The above results depict the behavior of both algorithms in depending on the \text{depth} of potential partial matches. The lines with square entry points represent problems where unification is easy. The lines with circle entry points are the same tests with hard unification problems.}%
  \label{fig:erase}
  \end{figure}

>>>>

With all of this in mind, in Fig.~\ref{fig:erase}, the test is set up as follows: given some rule $r$ with $n$ premises $P_i(x)$, $1 \leq i \leq n$, we run the procedure in a context where $n-1$ premises are instanciated with $a$ and one of them with $a + 1$.
The numbers on the $x$ axis represent which hypothesis has the incompatible instantiation in the order of consideration by the algorithm.
In a sense, this is how \textit{deep} the naive algorithm investigate before it stops.
The position of the incompatible hypothesis has little effect on the stateful algorithm as the index selects the rule in all case.
The work the stateful algorithm has to do is thus close to constant in this scenario.
The Figure \ref{fig:erase-noprecomp} showcases that the naive algorithm has a sizable advantage in this test, especially when unification is difficult.

@@

\begin{figure}
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.75]
      \begin{axis}[
        title={Depth Test},
        xlabel={Depth},
        ylabel={Time in ms},
        xmin=1, xmax=5,
        ymin=8, ymax=2^10,
        xtick={1,2,3,4,5},
        ymode=log,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 35.466292) (2, 40.918638) (3, 46.936125) (4, 54.933124) (5, 61.269861)
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 135.967736) (2, 127.324444) (3, 136.891250) (4, 118.720444) (5, 116.293222) 
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 107.978069) (2, 115.900527) (3, 122.454958) (4, 137.732764) (5, 144.037681) 
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 619.932611) (2, 601.456805) (3, 600.926110) (4, 589.144389) (5, 595.091180) 
          };
      \end{axis}
    \end{tikzpicture}
    \caption{Depth benchmark no-precompilation}%
    \label{fig:erase-noprecomp}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.75]
      \begin{axis}[
        title={Depth Test},
        xlabel={Depth},
        %ylabel={Time in ms},
        xmin=1, xmax=5,
        ymin=8, ymax=1024,
        xtick={1,2,3,4,5},
        ymode=log,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 13.728805) (2, 17.909194) (3, 23.324666) (4, 28.030819) (5, 33.279611) 
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 13.841305) (2, 12.801791) (3, 13.358389) (4, 13.200611) (5, 13.251902)
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 81.546486) (2, 85.345444) (3, 90.532347) (4, 98.247764) (5, 105.423875) 
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 25.539097) (2, 24.432485) (3, 25.629666) (4, 23.935250) (5, 23.911139)
          };
      \end{axis}
    \end{tikzpicture}
    % graph \includegraphics[width=.4\linewidth]{image1}
    \caption{Depth benchmark with precompilation}%
    \label{fig:erase-precomp}
  \end{subfigure}
  \caption{The above results depict the behavior of both algorithms in depending on the \text{depth} of potential partial matches. The lines with square entry points represent problems where unification is easy. The lines with circle entry points are the same tests with hard unification problems.}%
  \label{fig:erase}
  \end{figure}
